The Inner Mass-Ratio Distribution of Intermediate-Mass Stars

### 2016-01-28 paper/ms.tex
Added malmquist bias discussion to paper.

We are now finally in a position to estimate the mass-ratio distribution for our sample. We estimate the mass for each detected companion star by sampling the temperatures given in Table \ref{tab:companions} as a gaussian, and converting each temperature sample into a mass sample. We do the conversion to mass both by interpolating the spectral type relations from [CITE mamajek table], and by interpolating from temperature and system age (see Section \ref{sec:sp}) to mass with Dartmouth isochrones [CITE]. Both methods give similar results in most cases, although the isochrone masses are significantly higher for very young systems. We sample the mass ratio of the system by dividing the companion mass samples by samples of the primary mass (Section \ref{sec:sp}). We denote the $n$th mass ratio sample for the $k$th star as $q_k^{(n)}$, and denote the number of these samples as $N_k$.

We use the methodology developed in \citet{Foreman2014} to perform bayesian inference on the shape and form of the mass ratio distribution. The log-likelihood function in this formalism is:

\begin{multline}
\ln{\mathcal{L}(\{\vec{x_k}\}| \vec{\theta})} = -K \int_0^1 Q(q)P(q|\vec{\theta})dq + \\ \sum_{k=1}^K \ln{\frac{K}{N_k} \sum_{n=1}^{N_k} Q(q_k^{(n)}) P(q_k^{(n)}|\vec{\theta})}
\end{multline}

@@

\begin{align}
 P_1(q|\vec{\theta}) &= \begin{cases}
      \hfill \theta_1 \hfill & q \in \Delta_1 \\
      \hfill \theta_2 \hfill & q \in \Delta_2 \\
      \hfill \ldots \\
      \hfill \theta_7 \hfill & q \in \Delta_7
     \end{cases} \\
 P_2(q|\vec{\theta}) &= \frac{A}{q \sqrt{2\pi\sigma^2}} e^{-\frac{(\ln{q} - \mu)^2}{2\sigma^2}}
\end{align}
The constant $A$ in the log-normal distribution is a renormalization factor such that the distribution is only defined from $0 < q < 1$:

@@

\begin{equation}
P_1(\vec{\theta} | \{\vec{x_k}\}) \propto P_1(q|\vec{\theta}) P(\vec{\theta}| \alpha, m, \tau, \epsilon) P(\alpha, m, \tau, \epsilon)
\end{equation}

@@

We show the resulting distribution in Figure \ref{fig:mrd}. The $1\sigma$ uncertainties in the bin heights from the histogram model are shown as error bars, and we overplot 500 samples of the log-normal fit to show the spread allowed by the data.

>>>>

We are now finally in a position to estimate the mass-ratio distribution for our sample. We estimate the mass for each detected companion star by sampling the temperatures given in Table \ref{tab:companions} as a gaussian, and converting each temperature sample into a mass sample. We do the conversion to mass both by interpolating the spectral type relations from [CITE mamajek table], and by interpolating from temperature and system age (see Section \ref{sec:sp}) to mass with Dartmouth isochrones [CITE]. Both methods give similar results in most cases, although the isochrone masses are significantly higher for very young systems. We therefore only show the results obtained with the isochrone masses. We sample the mass ratio of the system by dividing the companion mass samples by samples of the primary mass (Section \ref{sec:sp}). We denote the $n$th mass ratio sample for the $k$th star as $q_k^{(n)}$, and denote the number of these samples as $N_k$.

\subsection{Fitting  Methodology}

We use the methodology developed in \citet{Foreman2014} to perform bayesian inference on the shape and form of the mass ratio distribution. The log-likelihood function in this formalism is derived from modeling the survey as a draw from the inhomogeneous Poisson process with rate density $\Gamma \equiv KQ(q)P(q)$:

\begin{multline}
\ln{\mathcal{L}(\{\vec{x_k}\}| \vec{\theta})} = -K \int_0^1 Q(q)P(q|\vec{\theta})dq + \\ \sum_{k=1}^K \ln{\frac{K}{N_k} \sum_{n=1}^{N_k} Q(q_k^{(n)}) P(q_k^{(n)}|\vec{\theta})}
\label{eqn:money}
\end{multline}

@@

\begin{align}
 P_1(q|\vec{\theta}) &= \begin{cases}
      \hfill \theta_1 \hfill & q \in \Delta_1 \\
      \hfill \theta_2 \hfill & q \in \Delta_2 \\
      \hfill \ldots \\
      \hfill \theta_7 \hfill & q \in \Delta_7
     \end{cases} \label{eqn:P1} \\
 P_2(q|\vec{\theta}) &= \frac{A}{q \sqrt{2\pi\sigma^2}} e^{-\frac{(\ln{q} - \mu)^2}{2\sigma^2}} \label{eqn:P2}
\end{align}
The constant $A$ in the log-normal distribution is a renormalization factor such that the distribution is only defined from $0 < q < 1$:

@@

\begin{equation}
P_1(\vec{\theta} | \{\vec{x_k}\}) \propto \mathcal{L}_1(\{\vec{x_k}\}| \vec{\theta}) P(\vec{\theta}| \alpha, m, \tau, \epsilon) P(\alpha, m, \tau, \epsilon)
\end{equation}

@@

\subsection{Malmquist Bias Correction}

We are trying to recover the intrinsic distribution from an observed sample; thus we should fit the data to the probability distribution function (PDF) for mass ratio, \emph{given that we observed the star}: $P(q|\vec{\theta}, \mathrm{obs})$. In a volume-limited sample, this is equal to $P(q|\vec{\theta})$; however, our sample is volume-limited and therefore suffers from Malmquist bias. There is a higher probability of equal-mass binary systems in our data because they contribute twice the flux and are therefore more likely to fall under the magnitude limit. We can calculate the PDF for mass ratio given that we observed the system from Bayes' theorem:

\begin{equation}
P(q|\vec{\theta}, \mathrm{obs}) = \frac{P(\mathrm{obs}|q, \vec{\theta}) P(q|\vec{\theta})}{\int_0^1 P(\mathrm{obs}|q, \vec{\theta}) P(q|\vec{\theta}) dq}
\end{equation}

We already know $P(q|\vec{\theta})$ (Equations \ref{eqn:P1} and \ref{eqn:P2}). We estimate $P(\mathrm{obs}|q)$ by simulating a very large sample of binary stars:

\begin{enumerate}
\item Sample primary star masses from the Kroupa IMF [CITE]
\item Sample a distance for each star from a disk with infinite extent and scale height of $300$ pc. 
\item For each $q$ from 0 to 1, in steps of 0.01:
\begin{enumerate}
  \item Add a companion star to each primary with the appropriate mass to make a binary system with mass ratio $q$.
  \item Calculate the combined absolute V-magnitude by interpolating the spectral type relations in [CITE mamajek]
  \item Calculate apparent magnitude $V$
  \item Find fraction of stars ($f(q)$) with apparent $V < 6$
\end{enumerate}
\item Fit the sampled fractions $f(q)$ to a 5th-order polynomial.
\end{enumerate}

With the fitted malmquist-correction polynomial, we then substitute $P(q|\vec{\theta}, \mathrm{obs})$ everywhere $P(q|\vec{\theta})$ appears in Equation \ref{eqn:money}.

We show the resulting fits in Figure \ref{fig:mrd}. The $1\sigma$ uncertainties in the bin heights from the histogram model are shown as error bars, and we overplot 500 samples of the log-normal fit to show the spread allowed by the data.
